### 2021-06-14 project_rodolfo/paper.tex
rodolfo anteproyecto

%\documentclass[final,3p]{article}
\documentclass[final,12p]{article}
\usepackage{graphicx}
\usepackage[letterpaper,margin=1.0in]{geometry}
\usepackage{amssymb}
\usepackage{changepage}
\usepackage{float}
\usepackage{hyperref}
\usepackage{url}
\usepackage{afterpage}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\usepackage[utf8]{inputenc} 
\usepackage{lastpage}

@@

%%%%%%%comands
\newcommand{\SubItem}[1]{ {\setlength\itemindent{15pt} \item[-] #1} }
\newcommand{\lumi}[1]{{#1~fb$^{-1}$}}
\newcommand{\instlumi}[1]{#1$\times 10^{34}$ cm$^{-2}$s$^{-1}$} }

@@

Luminosity, $L$, is a key parameter at particle colliders along with the energy available in the collision.
$L$ is one of the  main figures of merit that quantify the potential for observing new particles and measuring their properties.
The instantaneous luminosity $L(t)$ is the process-independent ratio between the rate $R(t)$ of events produced per unit time and the cross section for a given process $\sigma$: $L(t) = R(t)/\sigma$.
During Run 1 (2011-2012) LHC reached a peak instantaneous luminosity of \instlumi{0.77} and delivered an integrated luminosity of about \lumi{25} with a precision of about 2.0\% 
\footnote{1 barn is a unit of area corresponding to $10^{-24}$ cm${^2}$ and 1 femtobarn (fb) = $10^{-39}$ cm$^{2}$.
For comparison, the total Higgs production cross section is 48600 fb.}.
In the first part of Run 2 (2015-2016), the delivered luminosity has been measured to be \lumi{38.4} with an unprecedented precision of 1.3\% \cite{Sirunyan:2021qkt}.
For the second part of Run 2 (2017-2018), the integrated luminosity is about \lumi{78}, but its precise value and uncertainty remain to be determined and is one of the main objectives of this thesis project \cite{CMS:2018elu}.

The plan of the LHC till year 2038 is to obtain datasets with up to 10 times higher values of instanteneous luminosities in the final phase, the  High Luminosity LHC (HL-LHC), and corresponding integrated luminosities of order \lumi{3000} as shown in Figure~\ref{figure6}.
These final datasets will provide  precise measurements of the properties of the Higgs boson and other SM particles as shown in Fig. ~\ref{figureKappasUncs}.
This figure shows that one of the dominant uncertainties which remain to be determined is due to the luminosity measurement.
%In this project, we also perform studies for the design and construction of an upgraded luminosity measurement system to be installed for CMS in the HL-LHC phase\cite{Collaboration:2706512}.

@@

In the first part we propose to perform an analysis of the LHC luminosity data  recorded by the CMS experiment during 2017 and 2018 using Pixel Cluster Counting (PCC) method.

@@

The main objective of this work is to perform a measurement of the 2017-2018 LHC delivered luminosity with a precision at the level of 2\% or less using the CMS Pixel detector.
This precision is necesary for testing SM of elementary particles and identifying possible new physics which could elucidate the nature of Dark Matter in our universe.

%The second objective is to develop an upgraded luminosity measurement system for the HL-LHC phase using Tracker Endcap Pixel detector (TEPX). TEPX is a part of new tracker detector proposed by the CMS collaboration which will replace the current one during LS3 (2024-2026).

@@

As mentioned in the introduction, precision achieved in the measurement of integrated luminosity during Run 1 was about 2.5\% using  methods developed for the first time.
In 2015 and 2016 the uncertainty achieved was 1.6\% and 1.2\%, respectively, for datasets corresponding to \lumi{2.5} and \lumi{36}  \cite{Sirunyan:2021qkt}.
The methods used for the 2015 and 2016 data are similar to the strategy we propose in this project.
We expect that the final uncertainties for the 2017 and 2018 datasets to be similar, but the ultimate goal of the CMS collaboration is to reduce these uncertainties to below 1\%  \cite{Dainese:2019rgk}.

%For the upgrade part of this project, we must consider that in the HL-LHC phase the higher expected pileup (number of proton collisions per bunch crossing) will lead to more difficult scenarios for the luminosity measurement due to increased detector occupancy and larger irradiation of the detector components. 
%However, we expect that new methods to be introduced between now and the start of the HL-LHC phase will improve the measurement precision and compensate the higher backgrounds and detector aging.

@@

For the Run 2 luminosity measurement:
\begin{itemize}
\item understand the CMS pixel detector layout including the barrel layers and endcap disks and their constituent modules,
\item study the stability of the modules in the different layers or disks of the pixel detector and select the stable components,
\item calculate the visible cross section corresponding to the PCC rate for proton-proton colisions at 13 TeV using the selected pixel detector components,
\item study the stability and linearity of the PCC luminosity measurement by comparing to other CMS luminometers as a measurement of the systematic uncertainties,
\item determine the delivered luminosity per lumisection (23 second intervals) and the total integrated luminosity and determine the corresponding uncertainty.
\end{itemize}

@@

The PCC method consists of the reconstruction of track clusters produced by charged particle tracks as shown in Figure~\ref{fig:trackcluster}.
Due to the fine granularity of the pixel sensors and the large number of total pixels, the hit occupancy in the sensors remains very small, order of 1\%, during normal collisions.
This low occupancy makes the PCC  very linear as a function of pileup, an essential property of a good luminometer.

@@

For the Run 2 luminosity mesurement, the student will perform the studies for the selection of good modules for calculating luminosity using the PCC method.
From past performance of this algorithm it is expected that this measurement will give one of the best linearity and stability systematic uncertainties for the integrated luminosity value.
The results of these studies will be published in a peer-reviewed scientific journal.

Also, the student will become a part of an international experiment and is expected to present his results at national or international scientific conferences.

@@

\item {\bf Summer 2021}:
  \SubItem{ Readings on Standard Model of particle physics theory}
  \SubItem{ Readings on the LHC and CMS experiments}
  \SubItem{ Linux/Computing skills (\textsc{Bash, Emacs, c/c++, python, ROOT})}
  \SubItem{ Computing accounts at CERN and Fermilab}
  \SubItem{ Initial planning of the analysis strategy}

\item {\bf Semester 2021-2}:
  \SubItem{ Course I: on particle physics, particle detection, or data analysis}
  \SubItem{ Analysis/Computing skills (\textsc{Bash, Emacs, c/c++, python, ROOT})}
  \SubItem{ Analysis of the LHC Run 2 data using the PCC method}

\item {\bf Semester 2022-1}:
  \SubItem{ Completion of the analysis of the LHC Run 2 data using the PCC method}
  \SubItem{ Presentation of results at national or international conference}
  \SubItem{ Thesis writing}
  \SubItem{ Course II on particle physics, particle detection, or  data analysis}

>>>>

\documentclass[final,12pt]{article}
\usepackage{graphicx}
\usepackage[letterpaper,margin=1.0in]{geometry}
\usepackage{amssymb}
\usepackage{changepage}
\usepackage{float}
\usepackage{hyperref}
\usepackage{url}
\usepackage{afterpage}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\usepackage[utf8]{inputenc} 
\usepackage{lastpage}

@@

%%%%%%%comands
\newcommand{\SubItem}[1]{ {\setlength\itemindent{15pt} \item[-] #1} }
\newcommand{\lumi}[1]{{#1~fb$^{-1}$}}
\newcommand{\instlumi}[1]{#1$\times 10^{34}$ cm$^{-2}$s$^{-1}$}

@@

Luminosity, $L$, is a key parameter at particle colliders along with the energy available in the collision.
$L$ is one of the  main figures of merit that quantify the potential for observing new particles and measuring their properties.
The instantaneous luminosity $L(t)$ is the process-independent ratio between the rate $R(t)$ of events produced per unit time and the cross section for a given process $\sigma$: $L(t) = R(t)/\sigma$.
During Run 1 (2011-2012) LHC reached a peak instantaneous luminosity of \instlumi{0.77} and delivered an integrated luminosity of about \lumi{25} with a precision of about 2.0\% 
\footnote{1 barn is a unit of area corresponding to $10^{-24}$ cm${^2}$ and 1 femtobarn (fb) = $10^{-39}$ cm$^{2}$.
For comparison, the total Higgs production cross section is 48600 fb.}.
In the first part of Run 2 (2015-2016), the delivered luminosity has been measured to be \lumi{38.4} with an unprecedented precision of 1.3\% \cite{Sirunyan:2021qkt}.
For the second part of Run 2 (2017-2018), the integrated luminosity is about \lumi{78}, but its precise value and uncertainty remain to be determined and is the main objective of this thesis project \cite{CMS:2018elu}.

The plan of the LHC till year 2038 is to obtain datasets with up to 10 times higher values of instanteneous luminosities in the final phase, the  High Luminosity LHC (HL-LHC), and corresponding integrated luminosities of order \lumi{3000} as shown in Figure~\ref{figure6}.
These final datasets will provide  precise measurements of the properties of the Higgs boson and other SM particles as shown in Fig. ~\ref{figureKappasUncs}.
This figure shows that one of the dominant uncertainties which remain to be determined is due to the luminosity measurement.

@@

We propose to perform an analysis of the LHC luminosity data recorded by the CMS experiment during 2017 and 2018 using the Pixel Cluster Counting (PCC) method~\cite{LUM12001}.

@@

The main objective of this work is to perform a measurement of the 2017-2018 luminosity with a precision at the level of 2\% or less using the CMS Pixel detector.
This precision is necesary for testing SM of elementary particles and identifying possible new physics which could elucidate the nature of Dark Matter in our universe.

@@

As mentioned in the introduction, the precision achieved in the measurement of the integrated luminosity during Run 1 (2011-2012) was about 2.5\% using  methods developed for the first time.
In 2015 and 2016 runs, the uncertainty achieved was 1.6\% and 1.2\%, respectively, for datasets corresponding to \lumi{2.5} and \lumi{36}  \cite{Sirunyan:2021qkt}.
The methods used for the 2015 and 2016 data are similar to the strategy we propose in this project.
We expect that the final uncertainties for the 2017 and 2018 datasets to be similar, but the ultimate goal of the CMS collaboration is to reduce these uncertainties to below 1\%  \cite{Dainese:2019rgk}.

@@

\begin{itemize}
\item Understand the CMS pixel detector layout including the barrel layers and endcap disks and their constituent silicon pixel modules.
\item Study the stability of the modules in the different layers or disks of the pixel detector and select the stable components.
\item Determine the calibration constant corresponding to the PCC luminometer for proton-proton colisions at 13 TeV using the selected pixel detector components.
\item Determine the delivered luminosity per lumisection (23 second intervals) and the total integrated luminosity for 2017 or 2018 data.
\item Study some of the systematic uncertainties related to the measurement of the luminosity using the PCC. For example, calibration uncertainties, stability, and linearity.
\end{itemize}

@@

The PCC method consists of the reconstruction of track clusters produced by charged particle tracks as shown in Figure~\ref{fig:trackcluster}.
Due to the fine granularity of the pixel sensors and the large number of total pixels, the hit occupancy in the sensors remains very small (order of 1\%) during physics runs.
This low occupancy makes the PCC  very linear as a function of pileup, an essential property of a good luminometer.

@@

\begin{itemize}
\item Final selection of good modules for calculating luminosity using the PCC method.
\item Determination of the PCC calibration constant by analyzing the vdM scan data.
\item Determination of the luminosity per lumi section and integrated for the 2017 or 2018 data.
\item Estimation of the systematic uncertainties associated with the calibration constant or the integrated luminosity values.
\item A presentation of the results in a national or international conference.
\item Publication of the results in a peer-reviewed scientific journal.
\end{itemize}

@@

\item {\bf Summer 2021}:
  \SubItem{ Readings on Standard Model of particle physics,  the LHC and CMS experiments}
  \SubItem{ Linux/Computing skills (\textsc{Bash, Emacs, c/c++, python, ROOT})}
  \SubItem{ Computing accounts at CERN and Fermilab}
  \SubItem{ Initial planning of the analysis strategy}

\item {\bf Semester 2021-2}:
  \SubItem{ Course I: on particle physics, particle detection, or data analysis}
  \SubItem{ Analysis/Computing skills (\textsc{Bash, Emacs, c/c++, python, ROOT})}
  \SubItem{ Analysis of the LHC Run 2 luminosity data using the PCC method}

\item {\bf Semester 2022-1}:
  \SubItem{ Course II on particle physics, particle detection, or  data analysis}
  \SubItem{ Completion of the analysis of the LHC Run 2 luminosity data using the PCC method}
  \SubItem{ Presentation of results at national or international conference}
  \SubItem{ Thesis writing}
